\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-10 lang-design.tex
space hacks

@@

In an OO language, a subclass can override methods of a superclass.
Similar expressiveness is useful for metatheory mechanization, too.
%
For example, in a derived family, rather than adding new cases to an
induction proof, the programmer may prefer overriding the proof
entirely, as we observe in our case studies (\cref{sec:casestudies}).
%
Overriding is potentially incompatible with having equalities on late bound names, however.

Coq distinguishes \emph{opaque} definitions from \emph{transparent} ones.
\Lang supports the overriding of opaque definitions, which
include most proofs.
It is safe to override opaque definitions, because the type checker will
never attempt to unfold them.
For transparent definitions, the common case is that the programmer does
not want to override them.
In \cref{fig:stlc-mechanized}, "env", "empty", "subst", and "steps" are
transparent---that is, they are not defined with \lsti{Qed}.
\Lang treats transparent definitions as non-overridable by default.
Thus, the definitional equalities on "env", "empty", and "steps",
as well as the propositional equality on "subst", are available to the
type checker for type-checking the families.

@@

In \cref{fig:stlc-mechanized}, just because the "typesafe" theorem in
"STLC" is inherited and reused by "STLCFix", it does not follow that
in "STLCFix" the programmer can use "typesafe" to prove "progress".
If the programmer did, then they would be committing the logical fallacy
of circular reasoning: the proof of "progress" would depend on "typesafe",
yet the proof of "typesafe" depends on "progress".
%
Such circularity would easily lead to logical inconsistency:
%\YZ{Mention non-exhaustivity/overriding already poses challenges to ensuring consistency.}

@@

\noindent
In the program above, family~"B" extends~"A", overriding lemma~"f" by
proving it using lemma~"g".
Lemma~"g" is in turn inherited from family "A", where it is proven using
a late bound reference to lemma~"f".
Circularity between "f" and "g" allows proving \lsti{False}!

@@

Within family "STLCFix", a second partial recursor
(called "tm_prect_STLCFix") and its computational behaviors are axiomatized,
so properties such as \lsti[basicstyle=\fontsize{8.25}{9},columns=fixed,basewidth=0.9ex]{tm_fix x t1 = tm_fix x t2 ->  t1 = t2} can be proved.

>>>>

%\begin{centered}
%\begin{minipage}{.995\textwidth}
%\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
%\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
%tm_rect : \forall (P : tm -> Type), [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] ->
%  [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] -> [@\hlcase{tm-recursor-case-app}@](\forall t1, P t1 -> \forall t2, P t2 -> P (tm_app t1 t2))[@\tikzmarkend{tm-recursor-case-app}@] -> \forall t, P t
%\end{lstlisting}
%\end{minipage}
%\end{centered}

@@

In an OO language, a subclass can override methods of a superclass.
Similar expressivity is useful for mechanizing proofs, too.
%
For example, in a derived family, rather than adding new cases to an
induction proof, the programmer may prefer overriding the proof
entirely, as we observe in our case studies (\cref{sec:casestudies}).

Overriding is potentially incompatible with having equalities on late bound names, however.
%
Coq distinguishes \emph{opaque} definitions from \emph{transparent} ones.
\Lang supports the overriding of opaque definitions, which
include most proofs.
It is safe to override opaque definitions, because the type checker will
never attempt to unfold them.
For transparent definitions, the common case is that the programmer does
not want to override them.
In \cref{fig:stlc-mechanized}, "env", "empty", "subst", and "steps" are
transparent---that is, they are not defined with \lsti{Qed}.
\Lang treats transparent definitions as non-overridable by default.
Thus, the definitional equalities on "env", "empty", and "steps",
as well as the propositional equality on "subst", are available to the
type checker for type-checking the families.

@@

In \cref{fig:stlc-mechanized}, just because the "typesafe" theorem in
"STLC" is inherited and reused by "STLCFix", it does not follow that
in "STLCFix" the programmer can use "typesafe" to prove "progress".
If the programmer did, then they would be committing the logical fallacy
of circular reasoning: the proof of "progress" would depend on "typesafe",
yet the proof of "typesafe" depends on "progress".
%
Such circularity would easily lead to logical inconsistency.
Consider the following two families, where "B" extends "A":
%\YZ{Mention non-exhaustivity/overriding already poses challenges to ensuring consistency.}

@@

\noindent
"B" overrides lemma~"f" by proving it using~"g".
Lemma~"g" is in turn inherited from family "A", where it is proven using
a late bound reference to lemma~"f".
Circularity between "f" and "g" allows proving \lsti{False}!

@@

Within family "STLCFix", a second partial recursor
(called "tm_prect_STLCFix") and its computational behaviors are automatically axiomatized,
which allows properties of \lsti[basewidth=0.9ex]{tm_fix}, such as
\lsti[columns=fixed,basewidth=1.2ex]{tm_fix x t1 = tm_fix x t2 ->  t1 = t2}, to be proved.
